---
title: "Green turtle in southern CA Feb 2020"
output: html_notebook
---


```{r}
rm(list=ls())
library(lubridate)
library(RODBC)
library(tidyverse)
library(dplyr)
library(ggplot2)

SWFSC <- F
```

This file is used to extract green turtle sightings in southern CA for Megan's paper. Older data are extracted from the database directly. Newer data were typed into a spreadsheet, after finding out which data have not been entered using sightings_DB_update.Rmd.  New entries were made in data2upload_2020-02-06.xls. 

In order to do this extration at home, I saved the data into a local file. 

```{r}
if (SWFSC){
  turtle <- odbcConnect(dsn = 'Turtle', uid = '', pwd = '')

  # get SD Bay results:
  turtle.sightings <- sqlQuery(turtle, 'select * from tbl_Sighting') 
  
  odbcClose(turtle)
  
  turtle.sightings %>% select(Date_Observed,
                              Sighting_Number,
                              Species_ID,
                              City_ID,
                              County_ID,
                              State_ID,
                              Latitude,
                              Longitude) -> turtle.sightings
  
  readr::write_csv(turtle.sightings,
                   path = paste0("data/turtle_sightings.csv"))
  
  SWFSCcommon <- odbcConnect(dsn = "common", uid = "", pwd = "")
  
  # get the species table
  sp.table <- sqlQuery(SWFSCcommon, "select * from tblSpecies") %>% 
    select(-EditDate, -EditUserID, -RecordCreationDate, -NomenclatureAuthority, -Comments)
  county.table <- sqlQuery(SWFSCcommon, "select * from tblCounty") %>%
    select(-EditDate, -EditUserID, -RecordCreationDate)
  state.table <- sqlQuery(SWFSCcommon, "select * from tblState") %>%
    select(-EditDate, -EditUserID, -RecordCreationDate)
  
  city.table <- sqlQuery(SWFSCcommon, "select * from tblCity") %>%
    select(-EditDate, -EditUserID, -RecordCreationDate)
  
  odbcClose(SWFSCcommon)
  
  readr::write_csv(sp.table,
                   path = paste0("data/sp_table.csv"))
  
  readr::write_csv(county.table,
                   path = paste0("data/county_table.csv"))
  readr::write_csv(state.table,
                   path = paste0("data/state_table.csv"))
  readr::write_csv(city.table,
                   path = paste0("data/city_table.csv"))
  
} else {
  cols.sightings <- cols(Sighting_Number = col_factor(),
                         Species_ID = col_integer(),
                         City_ID = col_integer(),
                         County_ID = col_integer(),
                         State_ID = col_integer(),
                         Latitude = col_double(),
                         Longitude = col_double())
  
  turtle.sightings <- readr::read_csv(file = "data/turtle_sightings.csv",
                                      col_types = cols.sightings)
  
  cols.sp <- cols(ID = col_double(),
                  TaxanomicOrder = col_character(),
                  SubOrder = col_character(),
                  InfraOrder = col_character(),
                  Family = col_character(),
                  FamilyCommonName = col_character(),
                  Genus = col_character(),
                  Species = col_character(),
                  Subspecies = col_character(),
                  CommonName = col_character(),
                  SpType = col_character(),
                  SpName = col_character(),
                  CetaceanCode = col_double(),
                  AerialFisheryCode = col_character(),
                  BirderCode = col_character(),
                  AlternateCode1 = col_character(),
                  AlternateCode2 = col_character())
  
  sp.table <- readr::read_csv(file = "data/sp_table.csv",
                              col_types = cols.sp) 
  
  cols.county <- cols(ID = col_integer(),
                      Name = col_character(),
                      CountryID = col_integer(),
                      StateID = col_integer(),
                      OceanBasin = col_integer())
  county.table <- readr::read_csv(file = "data/county_table.csv",
                                  col_types = cols.county)
  
  cols.state <- cols(ID = col_integer(),
                     TwoLetterID = col_character(),
                     Name = col_character(),
                     CountryID = col_integer(),
                     OceanBasinID = col_integer())
  state.table <- readr::read_csv(file = "data/state_table.csv",
                                 col_types = cols.state)
  
  cols.city <- cols(ID = col_integer(),
                    Name = col_character(),
                    CountryID = col_integer(),
                    StateID = col_integer(),
                    CountyID = col_integer(),
                    OceanBasin = col_integer())
  city.table <- readr::read_csv(file = "data/city_table.csv",
                                col_types = cols.city)
}

```

Then, bring in the missing part - this is the part that was not entered into the database as of early February 2020. I have not been able to upload to the database so I have them in a spreadsheet. 

```{r}
#xlsx library requries Java... ITS doesn't like installing Java... 

new.data <- xlsx::read.xlsx(file = "data/data2upload_2020-02-06.xlsx",
                            sheetIndex = 1) %>% 
  select(Date_Observed,
         Sighting_Number,
         Species_ID,
         City_ID,
         County_ID,
         State_ID,
         Latitude,
         Longitude )

all.data <- rbind(turtle.sightings, new.data) 

# remove unknown species - this needs to be fixed eventually
all.data %>%
  filter(!is.na(Species_ID)) -> all.data.1

# prep county, city, species names
sp.table %>% mutate(Species_ID = ID) %>% #-> sp.table.2
  right_join(all.data.1, by = "Species_ID") %>% # -> sp.table.2
  select(Date_Observed, Sighting_Number, Species_ID, City_ID,
         County_ID, State_ID, Latitude, Longitude, Genus, Species) %>%
  filter(Genus == "Chelonia") -> green.sightings

county.table %>% mutate(County_ID = ID,
                        County_Name = Name) %>%
  select(County_ID, County_Name, StateID) %>%
  right_join(green.sightings, by = "County_ID") %>%
  filter(County_Name == "San Diego" |
           County_Name == "Orange" |
           County_Name == "Los Angeles" |
           County_Name == "Ventura" |
           County_Name == "Santa Barbara") %>%
  mutate(County_Name = factor(County_Name,
                              levels = c("Santa Barbara",
                                         "Ventura", 
                                         "Los Angeles",
                                         "Orange",
                                         "San Diego"))) -> green.SoCal.sightings

```


Let's make sure there are no duplicated entries.

```{r}
unique.sightings <- unique(green.SoCal.sightings$Sighting_Number)
length(unique.sightings) - nrow(green.SoCal.sightings)
```

Yes!!

Make a histogram
```{r}
p1 <- ggplot(data = green.SoCal.sightings) + 
  geom_bar(aes(x = County_Name))

p1
```

How about by year:

```{r}
green.SoCal.sightings %>% mutate(Year = year(Date_Observed)) -> green.SoCal.sightings

green.SoCal.sightings[is.na(green.SoCal.sightings$Year), "Year"] <- 2017

p2 <- ggplot(green.SoCal.sightings) + 
  geom_bar(aes(x = Year, fill = County_Name))
p2
```


Write out green results into a file:
```{r}
write.csv(green.SoCal.sightings,
          file = "data/SoCal_green_sightings.csv",
          quote = F, row.names = F)
```

