---
title: "Updating sightings database in a batch"
output: html_notebook
---

This script attempts to read sightings directories in a local hard drive and creates a text file that can be uploaded to the sightings database. The output of this script still needs to be populated prior to uploading. 

First initialize the workspace and load necessary libraries.
```{r}
library(lubridate)
library(RODBC)
library(tidyverse)
library(dplyr)

ifelse(Sys.info()[1] == 'Linux',
       source('~/Documents/R/tools/TomosFunctions.R'),
       source('~/R/tools/TomosFunctions.R'))

# The last run date of query - important only if local saved file is used - see below
run.date <- "2018-08-07"

# Whether or not at SWFSC - either access the database directory or use a saved file.
SWFSC <- T

# The location of all directories that are sightings specific
local.path <- "D:/TomosFolders/turtles/sightings/Turtle Sightings"
```

Then connect to the database:
```{r}
if (SWFSC){
  turtle <- odbcConnect(dsn = 'Turtle', uid = '', pwd = '')

  # get SD Bay results:
  turtle.sightings <- sqlQuery(turtle, 'select * from tbl_Sighting') 
  
    # select(NMFS_Tag, Turtle_ID, Year_caught,
    #        Month_caught, Day_caught, Caught_Dead,
    #        PIT_Tag_LFF, PIT_Tag_RFF, Inconel_Tag_LFF,
    #        Inconel_Tag_RFF, Sex, Weight_kg,
    #        Str_Carapace_Length_cm, Str_Carapace_Width_cm,
    #        Cur_Carapace_Length_cm, Cur_Carapace_Width_cm,
    #        Sat_Tag_ID, Satellite_Tag_Added) %>%
    # filter(., !is.na(NMFS_Tag)) %>%
    # filter(., !is.na(Sat_Tag_ID)) %>%
    # mutate(., Capture_Date = as.Date(paste0(Year_caught, '-',
    #                                         Month_caught, '-',
    #                                         Day_caught),
    #                                  format = '%Y-%m-%d'))

  odbcClose(turtle)
  # csv files don't work so well because descriptions contain commas... 
  # readr::write_csv(turtle.sightings, 
  #                  path = paste0('data/turtle_sightings_', 
  #                                Sys.Date(), ".csv"))
} else {
  turtle.sightings <- read.csv(paste0('data/turtle_sightings_', 
                                 run.date, ".csv"))
}
```

Then look at the directory in which all sightings data are stored in a local hard drive.

```{r}
local.dir <- list.dirs(path = local.path, recursive = FALSE)

# find all directory names, split them by the path name so that only sighting number
# and name of the observer are kept. Then, split those into two fields and
# name them correctly
sightings.df <- apply(as.array(local.dir), 
                      FUN = function(x) unlist(strsplit(x, 
                                                        split = paste0(local.path,
                                                                       "/")))[2], 
                      MARGIN = 1) %>%
  strsplit(" - ") %>% unlist() %>%
  matrix(ncol = 2, byrow = TRUE) %>%
  as.data.frame(stringsAsFactors = FALSE) %>% 
  setNames(c("Sighting_Number", "Observed_By"))

```

Now that we have the most current database entries and what we have in the hard drive, we do an anti-join to pick up those from the hard drive that are not in the database. Then, create a new dataframe with all necessary columns, fill in the two columns, and save it as a text file.

```{r}
# Warning returns because Sighting_Number in turtle.sightings is a factor variable.
missing.sightings <- anti_join(sightings.df, 
                               turtle.sightings, 
                               by = "Sighting_Number")

sighting.date <- apply(missing.sightings, 
                       FUN = function(x) unlist(strsplit(x, split = "-"))[1], 
                       MARGIN = 1) %>%
  as.numeric() %>%
  yyyymmdd2date()

col.names <- colnames(turtle.sightings)
out.df <- setNames(data.frame(matrix(ncol = length(col.names),
                                     nrow = nrow(missing.sightings))), 
                   col.names)

out.df[, "Sighting_Number"] <- missing.sightings$Sighting_Number
out.df[, "Observed_By"] <- missing.sightings$Observed_By
out.df[, "Date_Observed"] <- sighting.date

# write.csv(out.df, 
#           file = paste0("data/data2upload_", Sys.Date(), ".csv"),
#           quote = FALSE, na = "")
```
